---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  eval=TRUE
)
```
# serosolver

[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)

`serosolver` uses a hierarchical model with a custom Markov chain Monte Carlo sampler to simultaneously infer antibody kinetics and infection histories from cross-sectional or longitudinal serological data. `serosolver` is a [time-since-infection serodynamics model](https://osf.io/preprints/osf/kqdsn), meaning that infection times are back-calculated from one or more antibody measurements through an antibody kinetics model. `serosolver` can be used to infer infection timings during a study period using longitudinal measurements against a single antigen, or lifetime infection histories using multi-antigen serology panels. The package and model are described by Hay _et al._ [here](https://doi.org/10.1371/journal.pcbi.1007840).

## Recent changes
`serosolver` is back in active development to fix bugs, standardize variable names and add new features.
<details>
  <summary>List of recent changes:</summary>
  
* Overhaul of variable names (e.g., _titre_ -> _measurement_, _strain_ -> _biomarker_id_)
* Consolidation of plotting functions
* Moving more options and inputs behind the scenes to streamline the user interface
* Generalization to consider multiple biomarker types per sample (e.g., antibody titre and avidity)
* Support for continuous as well as discrete observations (e.g., can now fit to ELISA data as well as HAI titres)
* _IN PROGRESS_ Some small improvements to the MCMC sampler and parameter transformations
* _IN PROGRESS_ Improved guidance and support for using priors
* _IN PROGRESS_ Model infection histories and antibody kinetics as a function of demographic variables
* _IN PROGRESS_ Allow some infection states to be fixed during fitting
* _IN PROGRESS_ Inclusion of explicit immunity model
* _IN PROGRESS_ Ways to fix or estimate starting/baseline titres
* _IN PROGRESS_ Added tests
</details>

## Installation
1. Install [R][r-project]

2. Install the development version of serosolver from [GitHub](https://github.com/seroanalytics/serosolver):

```{r installation,eval=FALSE}
devtools::install_github("seroanalytics/serosolver")
library(serosolver)
```
## Quick start and vignettes

Read the [quick start vignette][vignette-doc] to set up and run a simple implementation with a simulation model.

Additional vignettes:

* [Longitudinal data][c1-vignette-doc]: estimating infection timings using longitudinal data, example of influenza A/H1N1p in Hong Kong
* [Cross-sectional data][c2-vignette-doc]: estimating life-course infection histories from multi-strain serology, example of influenza A/H3N2 from the [Fluscape cohort](https://pubmed.ncbi.nlm.nih.gov/26875566/)
* [Overview of optional features]: walkthrough of additional `serosolver` features and use cases, such as inclusion of biomarker-specific measurement offsets
* [Multiple measurements]: fitting `serosolver` to multiple biomarker types, example of binding avidity and ELISA measurements per sample
* [Group-level differences]: estimating demographic differences in antibody kinetics and attack rates

[r-project]: http://cran.r-project.org
[vignette-doc]: https://seroanalytics.github.io/serosolver/articles/serosolver-quick_start_guide.html
[c1-vignette-doc]: https://seroanalytics.github.io/serosolver/articles/cs1_vignette.html
[c2-vignette-doc]: https://seroanalytics.github.io/serosolver/articles/cs2_vignette.html

## Example

This is a basic example of simulating some serological data and fitting the model using the MCMC framework.

```{r example_setup, message=FALSE,warning=FALSE}
library(serosolver)
library(ggplot2)
library(plyr)
library(dplyr)
library(tidyr)
library(data.table)
library(doParallel)
library(coda)

## Load in example parameter values and antigenic map
data(example_par_tab)
data(example_antigenic_map)
data(example_antibody_data)
data(example_inf_hist)
```
```{r example_plot, message=FALSE,warning=FALSE}
plot_antibody_data(example_antibody_data,example_antigenic_map$inf_times,n_indivs=1:4,example_inf_hist)
```

```{r example_mcmc, message=FALSE,warning=FALSE}
## Run the MCMC
# This example uses prior version 2 (i.e. beta prior on phi with parameters shape1 and shape2)
output <- serosolver::serosolver(example_par_tab, example_antibody_data, example_antigenic_map,
                filename="readme", prior_version=2,n_chains=3,parallel=TRUE,
                mcmc_pars=c(adaptive_iterations=10000, iterations=50000),verbose=FALSE)
```
```{r example_model_fits, message=FALSE,warning=FALSE}
# Plot model predicted titres for a subset of individuals
chains <- load_mcmc_chains(location=getwd(),par_tab=example_par_tab,burnin = 10000,unfixed=TRUE)
plot_model_fits(chain = chains$theta_chain,
                infection_histories = chains$inf_chain,
                known_infection_history = example_inf_hist,
                antibody_data = example_antibody_data,individuals=1:4,
                antigenic_map=example_antigenic_map,
                par_tab=example_par_tab,
                orientation="cross-sectional")

```


