---
title: "Advanced features"
subtitle: Using additional model controls in `serosolver`
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
editor_options:
  chunk_output_type: console
bibliography: vignette_references.bib
vignette: |
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{Advanced features}
  %\VignetteEngine{knitr::rmarkdown}
---

<!-- Modified by an AI assistant on 2026-09-17 using GPT-5. Updated worked
observation-model examples to use character data_type labels. -->
<!-- Modified by an AI assistant on 2026-09-17 using GPT-5. Completed a consistency pass over the advanced-feature workflows, including the offset comparison, prior-version-1 outputs, and biomarker-group labels. -->

```{r setup, include=FALSE}
set.seed(1)
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", eval = TRUE,
  fig.width = 8, fig.height = 6, dpi = 300, out.width = "100%"
)
options(digits = 3)

library(serosolver)
library(ggplot2)
library(dplyr)
library(tidyr)
#vignette_dir <- dirname("vignettes/advanced_features.Rmd")
vignette_dir <- dirname(knitr::current_input())
cache_root <- file.path(
  vignette_dir, "..", "inst", "extdata", "advanced_features"
)

## Used only while knitting so that long fits can be reused without changing
## the workflow shown to the reader.
fit_or_load <- function(par_tab, antibody_data, cache_folder, file_stem,
                        mcmc_pars, ...) {
  cache_dir <- file.path(cache_root, cache_folder)
  filename <- file.path(cache_dir, file_stem)
  dir.create(cache_dir, recursive = TRUE, showWarnings = FALSE)
  chain_files <- list.files(
    cache_dir,
    pattern = paste0("^", file_stem, "_[0-9]+_chain\\.csv$"),
    full.names = TRUE
  )
  history_files <- sub("_chain\\.csv$", "_infection_histories.csv", chain_files)
  settings_file <- paste0(filename, "_serosolver_settings.RData")
  cache_available <- length(chain_files) == 3L &&
    file.exists(settings_file) && all(file.exists(history_files))

  if (!cache_available) {
    return(do.call(
      serosolver,
      c(
        list(
          par_tab = par_tab, antibody_data = antibody_data,
          filename = filename, mcmc_pars = mcmc_pars,
          n_chains = 3, parallel = TRUE, plot_outputs = TRUE,
          verbose = FALSE
        ),
        list(...)
      )
    ))
  }

  load(settings_file)
  chains <- load_mcmc_chains(
    location = cache_dir, par_tab = par_tab,
    burnin = mcmc_pars[["adaptive_iterations"]],
    thin = 1, convert_mcmc = FALSE, verbose = FALSE
  )
  diagnostics <- suppressMessages(plot_mcmc_diagnostics(
    location = cache_dir, par_tab = par_tab,
    burnin = mcmc_pars[["adaptive_iterations"]]
  ))
  list(all_diagnostics = diagnostics, settings = serosolver_settings,
       mcmc_chains = chains)
}

mcmc_pars <- c(
  adaptive_iterations = 5000, iterations = 100000,
  thin = 200, thin_inf_hist = 1000, save_block = 1000
)
```

# Overview

The main [serosolver guide](serosolver-guide.html) introduces the usual `serosolver` workflow, including output files and convergence diagnostics. 
In practice, however, the available data do not always
fit that simple setup. An infection may be known from another source, an
individual may already have antibody at the start of the analysis, or
different assays may measure related but distinct antibody responses. 

`serosolver` supports a number of advanced features to accommodate more complex systems and assumptions. These include:

1. Fixing known infection states
2. Setting starting antibody levels rather than building antibody profiles from zero
3. Accounting for systematic measurement biases between biomarker IDs
4. Alternative antibody kinetics model forms (e.g., exponential waning)
5. Multiple observation types per sample (referred to as biomarker groups), such as antibody titres and antibody avidity
6. An alternative way of parameterising the prior version on the infection histories, giving more flexibility in the assumed epidemiological dynamics

# Shared synthetic data

This vignette is based around a simulated, small longitudinal serosurvey. We assume there are 120 individuals,
five serum samples per person, and 25 possible infection times. Antibody is
measured against a subset of the possible biomarker IDs.

```{r shared-data}
set.seed(123)
possible_exposure_times <- 1:25
sampling_times <- c(8, 12, 16, 20, 24)
measured_biomarker_ids <- seq(1, 25, by = 2)
n_indiv <- 120

antigenic_map <- data.frame(
  x_coord = seq(0, 4, length.out = length(possible_exposure_times)),
  y_coord = sin(seq(0, 2 * pi, length.out = length(possible_exposure_times))),
  inf_times = possible_exposure_times
)
demographics <- data.frame(
  individual = seq_len(n_indiv),
  birth = sample(1:4, n_indiv, replace = TRUE)
)

data(example_par_tab)
base_par_tab <- example_par_tab %>% filter(names != "phi")
base_par_tab$values[base_par_tab$names == "min_measurement"] <- 0
base_par_tab$values[base_par_tab$names == "max_measurement"] <- 10
base_par_tab$lower_bound[base_par_tab$names == "min_measurement"] <- 0
base_par_tab$upper_bound[base_par_tab$names == "min_measurement"] <- 0
base_par_tab$lower_bound[base_par_tab$names == "max_measurement"] <- 10
base_par_tab$upper_bound[base_par_tab$names == "max_measurement"] <- 10

attack_rates <- simulate_attack_rates(
  infection_years = possible_exposure_times,
  mean_par = 0.15, sd_par = 0.05, n_groups = 1
)
simulated_data <- simulate_data(
  par_tab = base_par_tab, n_indiv = n_indiv,
  possible_exposure_times = possible_exposure_times,
  measured_biomarker_ids = measured_biomarker_ids,
  sampling_times = sampling_times, nsamps = length(sampling_times),
  antigenic_map = antigenic_map, attack_rates = attack_rates,
  data_type = "discrete", demographics = demographics
)
antibody_data <- as.data.frame(simulated_data$antibody_data)
head(check_data(antibody_data))
```

# Fixed infection states

In the usual workflow, all infection states are estimated from the serological
data. Sometimes there is additional information about a particular individual
and time period, such as a confirmed infection or a period during which
infection was known not to have occurred. These states can be supplied as a
long data frame with `individual`, `time`, and `value` columns. The supplied
rows are held fixed while the remaining infection states are estimated.

Here, 10% of the simulated infection-state entries are treated as known.

```{r fixed-states-data}
infection_matrix <- as.matrix(simulated_data$infection_histories)
all_infection_states <- data.frame(
  individual = rep(seq_len(nrow(infection_matrix)),
                   each = length(possible_exposure_times)),
  time = rep(possible_exposure_times, times = nrow(infection_matrix)),
  value = as.vector(t(infection_matrix))
)
set.seed(123)
known_state_rows <- sample(
  seq_len(nrow(all_infection_states)),
  size = ceiling(0.10 * nrow(all_infection_states))
)
fixed_inf_hists <- all_infection_states[known_state_rows, ]
head(fixed_inf_hists)
```

```{r fixed-states-call, eval=FALSE}
fixed_fit <- serosolver(
  par_tab = base_par_tab, antibody_data = antibody_data,
  antigenic_map = antigenic_map,
  possible_exposure_times = possible_exposure_times,
  fixed_inf_hists = fixed_inf_hists,
  filename = "fixed_infection_states", data_type = "discrete",
  n_chains = 3, parallel = TRUE, mcmc_pars = mcmc_pars
)
```

```{r fixed-states-fit, include=FALSE}
fixed_fit <- fit_or_load(
  par_tab = base_par_tab, antibody_data = antibody_data,
  cache_folder = "fixed_infection_states", file_stem = "fixed_infection_states",
  mcmc_pars = mcmc_pars, antigenic_map = antigenic_map,
  possible_exposure_times = possible_exposure_times,
  fixed_inf_hists = fixed_inf_hists, data_type = "discrete"
)
```

The model-fit plot uses the full simulated history as a reference. Only the
10% selected above were supplied to the fit; the remaining states were still
estimated from the antibody data. In the infection-probability panel below,
the blue points mark some of the supplied fixed states. A fixed infection has
posterior probability 1 at that time, while a fixed non-infection has
posterior probability 0.

```{r fixed-states-output}
fixed_fit$all_diagnostics$theta_estimates %>%
  select(names, median, lower95_CrI, upper95_CrI) %>% head(10)

plot_model_fits(
  fixed_fit$mcmc_chains$theta_chain, fixed_fit$mcmc_chains$inf_chain,
  settings = fixed_fit$settings, individuals = 1:4,
  known_infection_history = infection_matrix,
  orientation = "cross-sectional", expand_to_all_times = TRUE
)[[1]] + theme(axis.text.x = element_text(angle = 45, hjust = 1))

fixed_states_to_plot <- fixed_inf_hists %>%
  filter(individual %in% 1:16) %>%
  transmute(i = individual, j = time, value = value)

plot_cumulative_infection_histories(
  fixed_fit$mcmc_chains$inf_chain,
  indivs = 1:16,
  real_inf_hist = infection_matrix,
  possible_exposure_times = possible_exposure_times,
  pad_chain = TRUE
)[[2]] +
  geom_point(
    data = fixed_states_to_plot,
    aes(x = j, y = value),
    inherit.aes = FALSE,
    shape = 21, fill = "white", colour = "blue", size = 2
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# Starting antibody levels

The antibody model needs a level from which to start solving each
individual's trajectory. By default, these starting levels are set to zero, assuming that individual's begin life with no detectable antibodies (assuming that all maternal antibodies have waned).
That is sensible when the first possible infection time is early enough that
pre-existing antibody can be ignored, but it may be less suitable when the
analysis begins part way through an individual's antibody history.

There are several ways to provide starting levels. The short character options
are useful for simpler analyses, while an explicit data frame gives the most
control. The following code records the default and median-based options. We
then create a data set whose antibody levels are non-zero at the start of the
observation window and vary between individuals, while retaining the usual
simulated infection histories. To keep this example focused on starting
levels, it uses one antigenic coordinate and one measured biomarker ID.

```{r starting-level-options}
starting_levels_default <- "none"
starting_levels_median <- "median"
starting_level_function <- function(data) {
  create_start_level_data(data, start_level_summary = "none", randomize = FALSE) %>%
    mutate(starting_level = 2 + (individual - 1) %% 4)
}

starting_antigenic_map <- data.frame(
  x_coord = 0, y_coord = 0, inf_times = possible_exposure_times
)
starting_sampling_times <- sampling_times
starting_attack_rates <- attack_rates
set.seed(124)
starting_simulated_data <- simulate_data(
  par_tab = base_par_tab, n_indiv = n_indiv,
  possible_exposure_times = possible_exposure_times,
  measured_biomarker_ids = max(possible_exposure_times),
  sampling_times = starting_sampling_times, nsamps = length(starting_sampling_times),
  antigenic_map = starting_antigenic_map, attack_rates = starting_attack_rates,
  data_type = "discrete", demographics = demographics,
  starting_levels = starting_level_function
)
starting_antibody_data <- as.data.frame(starting_simulated_data$antibody_data)
starting_infection_matrix <- as.matrix(starting_simulated_data$infection_histories)
starting_levels_explicit <- starting_simulated_data$start_levels
head(starting_levels_explicit)
```

```{r starting-level-call, eval=FALSE}
starting_fit <- serosolver(
  par_tab = base_par_tab, antibody_data = starting_antibody_data,
  antigenic_map = starting_antigenic_map,
  possible_exposure_times = possible_exposure_times,
  start_level = starting_levels_explicit,
  filename = "starting_levels_one_biomarker", data_type = "discrete",
  n_chains = 3, parallel = TRUE, mcmc_pars = mcmc_pars
)
```

The shorter alternatives would be passed as `start_level = "none"` (i.e., ignore starting levels) or
`start_level = "median"` (i.e., use the median of the earliest measurements for that individual against that biomarker ID). 
An explicit table is useful when starting levels come from another analysis or need to be set differently for particular
individuals, biomarker IDs, or biomarker groups. More details on each option can be found in the help files 

```{r starting-level-fit, include=FALSE}
starting_fit <- fit_or_load(
  par_tab = base_par_tab, antibody_data = starting_antibody_data,
  cache_folder = "starting_levels", file_stem = "starting_levels_one_biomarker",
  mcmc_pars = mcmc_pars, antigenic_map = starting_antigenic_map,
  possible_exposure_times = possible_exposure_times,
  start_level = starting_levels_explicit, data_type = "discrete"
)
```

Starting antibody levels represent antibody present before the first modelled
infection time. Here, the observations are generated using both
individual-specific starting levels and simulated infection histories. The
model-fit plot compares fitted trajectories with these observations, and the
infection-history plot shows the simulated infection histories.

```{r starting-level-output}
starting_levels_plot <- starting_levels_explicit %>%
  filter(individual %in% 1:4) %>%
  select(individual, biomarker_id, biomarker_group, starting_level) %>%
  distinct() %>%
  arrange(individual, biomarker_group, biomarker_id) %>%
  mutate(start_index = row_number())

plot_model_fits(
  starting_fit$mcmc_chains$theta_chain,
  starting_fit$mcmc_chains$inf_chain,
  antibody_data = starting_antibody_data,
  demographics = demographics,
  individuals = 1:4,
  par_tab = base_par_tab,
  antigenic_map = starting_antigenic_map,
  possible_exposure_times = possible_exposure_times,
  known_infection_history = starting_infection_matrix,
  start_level = starting_levels_plot,
  data_type = "discrete",
  orientation = "longitudinal",
  expand_to_all_times = TRUE
)[[1]]
```

# Measurement offsets

Measurements against different biomarkers can have systematic shifts that are
not explained by the shared antibody kinetics model. A measurement offset is
an estimated additive shift for a particular biomarker and observation group.
The parameter is called `rho`. It is added to the model-predicted antibody
level for the relevant measurements before the observation model is applied.

`add_rhos_par_tab()` appends one `rho` row to `par_tab` for each measured
biomarker and observation group. These rows have `par_type = 3`. The `par_tab` rows do not
contain the biomarker IDs themselves, so the accompanying
`measurement_indices` table provides the link: `rho_index = 1` refers to the
first `rho` row, `rho_index = 2` to the second, and so on.

```{r measurement-bias-setup}
offset_setup <- add_rhos_par_tab(
  base_par_tab, sampled_viruses = measured_biomarker_ids, n_obs_types = 1
)
offset_par_tab <- offset_setup[[1]]
measurement_indices <- offset_setup[[2]]
offset_par_tab$values[offset_par_tab$names == "rho"] <-
  seq(-2, 2, length.out = length(measured_biomarker_ids))
offset_par_tab %>%
  filter(names == "rho") %>%
  select(names, par_type, biomarker_group, values, lower_bound, upper_bound)
head(measurement_indices)
```

The simulated data below include different known shifts for the measured
biomarkers. The fitted model receives the `measurement_indices` table but is
not given the values of the shifts; it estimates the `rho` parameters from the
data. In practice, `rho` estimates are often highly correlated with attack-rate
estimates because both can explain higher observed antibody levels. This
feature therefore works best when the `rho` parameters are fixed at known or
assumed values, unless the data provide enough information to separate the two
effects.

```{r measurement-bias-call, eval=FALSE}
offset_fit <- serosolver(
  par_tab = offset_par_tab, antibody_data = offset_data,
  antigenic_map = antigenic_map,
  possible_exposure_times = possible_exposure_times,
  measurement_bias = measurement_indices,
  filename = "measurement_bias", data_type = "discrete",
  n_chains = 3, parallel = TRUE, mcmc_pars = mcmc_pars
)
```

```{r measurement-bias-data-fit, include=FALSE}
set.seed(124)
offset_simulated <- simulate_data(
  par_tab = offset_par_tab, n_indiv = n_indiv,
  possible_exposure_times = possible_exposure_times,
  measured_biomarker_ids = measured_biomarker_ids,
  sampling_times = sampling_times, nsamps = length(sampling_times),
  antigenic_map = antigenic_map, attack_rates = attack_rates,
  data_type = "discrete", demographics = demographics,
  measurement_bias = measurement_indices
)
offset_data <- as.data.frame(offset_simulated$antibody_data)
set.seed(124)
unshifted_simulated <- simulate_data(
  par_tab = base_par_tab, n_indiv = n_indiv,
  possible_exposure_times = possible_exposure_times,
  measured_biomarker_ids = measured_biomarker_ids,
  sampling_times = sampling_times, nsamps = length(sampling_times),
  antigenic_map = antigenic_map, attack_rates = attack_rates,
  data_type = "discrete", demographics = demographics
)
unshifted_data <- as.data.frame(unshifted_simulated$antibody_data)
offset_fit <- fit_or_load(
  par_tab = offset_par_tab, antibody_data = offset_data,
  cache_folder = "measurement_bias", file_stem = "measurement_bias",
  mcmc_pars = mcmc_pars, antigenic_map = antigenic_map,
  possible_exposure_times = possible_exposure_times,
  measurement_bias = measurement_indices, data_type = "discrete"
)
```

The histograms below compare observations generated from the same simulated
infection histories, with and without the measurement offsets. Three biomarker
IDs are shown to make the shifts easy to see.

```{r measurement-bias-simulation-plot}
comparison_ids <- measured_biomarker_ids[
  c(1, ceiling(length(measured_biomarker_ids) / 2), length(measured_biomarker_ids))
]
offset_comparison <- bind_rows(
  offset_data %>% mutate(offset = "With measurement offsets"),
  unshifted_data %>% mutate(offset = "Without measurement offsets")
) %>%
  filter(biomarker_id %in% comparison_ids)

ggplot(offset_comparison, aes(x = measurement, fill = offset)) +
  geom_histogram(binwidth = 1, boundary = 0, position = "identity", alpha = 0.5) +
  facet_wrap(~biomarker_id) +
  labs(x = "Observed measurement", y = "Number of observations", fill = NULL) +
  theme_bw()
```

The point ranges below show the posterior medians and 95% credible intervals
for the `rho` parameters. The open points show the values used to simulate the
data. Repeated `rho` rows are given unique names when the chains are read, so
they appear as `rho`, `rho.1`, `rho.2`, and so on.

```{r measurement-bias-output,fig.width=7,fig.height=4}
rho_estimates <- offset_fit$all_diagnostics$theta_estimates %>%
  filter(grepl("^rho(\\.[0-9]+)?$", names)) %>%
  select(names, median, lower95_CrI, upper95_CrI)
rho_truth <- offset_par_tab %>%
  filter(names == "rho") %>%
  transmute(names = make.unique(as.character(names)), true = values)
rho_plot_data <- left_join(rho_estimates, rho_truth, by = "names") %>%
  mutate(names = factor(names, levels = names))

ggplot(rho_plot_data, aes(x = names, y = median)) +
  geom_pointrange(
    aes(ymin = lower95_CrI, ymax = upper95_CrI,
        colour = "Posterior median and 95% CrI")
  ) +
  geom_point(
    aes(y = true, shape = "True simulated rho"),
    colour = "black", fill = "white", size = 2.5
  ) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
  scale_colour_manual(name = NULL, values = c("Posterior median and 95% CrI" = "#0072B2")) +
  scale_shape_manual(name = NULL, values = c("True simulated rho" = 21)) +
  labs(x = "Measurement offset parameter", y = "rho") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# Exponential waning

The standard model assumes linear waning of log antibody levels, which means that the remaining boost
falls by a fixed proportion per unit of time until it reaches zero.
We have found that some data are better described by assuming exponential waning of log antibody levels. This can be achieved by setting
`exponential_waning = TRUE`, which uses this exponential form for the short- and
long-term responses. It also changes the cross-reactivity model from a linear
decline with antigenic distance to an exponential decline.

```{r exponential-call, eval=FALSE}
exponential_fit <- serosolver(
  par_tab = base_par_tab, antibody_data = exponential_data,
  antigenic_map = antigenic_map,
  possible_exposure_times = possible_exposure_times,
  exponential_waning = TRUE,
  filename = "exponential_waning", data_type = "discrete",
  n_chains = 3, parallel = TRUE, mcmc_pars = mcmc_pars
)
```

```{r exponential-fit, include=FALSE}
exponential_simulated <- simulate_data(
  par_tab = base_par_tab, n_indiv = n_indiv,
  possible_exposure_times = possible_exposure_times,
  measured_biomarker_ids = measured_biomarker_ids,
  sampling_times = sampling_times, nsamps = length(sampling_times),
  antigenic_map = antigenic_map, attack_rates = attack_rates,
  data_type = "discrete", demographics = demographics, exponential_waning = TRUE
)
exponential_data <- as.data.frame(exponential_simulated$antibody_data)
exponential_fit <- fit_or_load(
  par_tab = base_par_tab, antibody_data = exponential_data,
  cache_folder = "exponential_waning", file_stem = "exponential_waning",
  mcmc_pars = mcmc_pars, antigenic_map = antigenic_map,
  possible_exposure_times = possible_exposure_times,
  exponential_waning = TRUE, data_type = "discrete"
)
```

```{r exponential-output,fig.height=6,fig.width=8}
exponential_fit$all_diagnostics$theta_estimates %>%
  filter(names %in% c("wane_short", "wane_long", "cr_short", "cr_long")) %>%
  select(names, median, lower95_CrI, upper95_CrI)

plot_estimated_antibody_model(
  exponential_fit$mcmc_chains$theta_chain, settings = exponential_fit$settings,
  solve_times = possible_exposure_times
) + coord_cartesian(ylim = c(0, 10))
```

# Multiple biomarker groups

A study may measure more than one antibody outcome for each individual, such
as a discrete antibody titre and a continuous avidity measurement. These are
represented as separate `biomarker_group`s. Each group has its own observation
model and antibody-kinetics parameters, but the groups share the same
individual infection histories. The two measurements can therefore provide
complementary information about the same infections.

```{r multiple-biomarker-setup}
multi_par_tab <- extend_par_tab_biomarker_groups(base_par_tab, 2)
multi_par_tab$values[multi_par_tab$biomarker_group == 2 &
                     multi_par_tab$names == "boost_long"] <- 5.5
multi_par_tab$values[multi_par_tab$biomarker_group == 2 &
                     multi_par_tab$names == "boost_short"] <- 4.5
multi_par_tab$values[multi_par_tab$biomarker_group == 2 &
                     multi_par_tab$names == "boost_delay"] <- 1.5
multi_par_tab$values[multi_par_tab$biomarker_group == 2 &
                     multi_par_tab$names == "wane_long"] <- 0.08
multi_par_tab$values[multi_par_tab$biomarker_group == 2 &
                     multi_par_tab$names == "wane_short"] <- 0.6
multi_par_tab$values[multi_par_tab$biomarker_group == 2 &
                     multi_par_tab$names == "cr_long"] <- 0.25
multi_par_tab$values[multi_par_tab$biomarker_group == 2 &
                     multi_par_tab$names == "cr_short"] <- 0.75
multi_antigenic_map <- bind_rows(
  antigenic_map %>% mutate(biomarker_group = 1),
  antigenic_map %>% mutate(biomarker_group = 2)
)
multi_simulated <- simulate_data(
  par_tab = multi_par_tab, n_indiv = n_indiv,
  possible_exposure_times = possible_exposure_times,
  measured_biomarker_ids = measured_biomarker_ids,
  sampling_times = sampling_times, nsamps = length(sampling_times),
  antigenic_map = multi_antigenic_map, attack_rates = attack_rates,
  data_type = c("discrete", "continuous"), demographics = demographics
)
multi_data <- as.data.frame(multi_simulated$antibody_data)
table(multi_data$biomarker_group)
```

The first observation group below is discrete and bounded, while the second is
continuous and bounded. The distinction is specified by the corresponding
entries in `data_type`; both groups still contribute to the same
infection-history inference.

```{r multiple-biomarker-call, eval=FALSE}
multi_fit <- serosolver(
  par_tab = multi_par_tab, antibody_data = multi_data,
  antigenic_map = multi_antigenic_map,
  possible_exposure_times = possible_exposure_times,
  data_type = c("discrete", "continuous"), filename = "multiple_biomarker_groups",
  n_chains = 3, parallel = TRUE, mcmc_pars = mcmc_pars
)
```

```{r multiple-biomarker-fit, include=FALSE}
multi_fit <- fit_or_load(
  par_tab = multi_par_tab, antibody_data = multi_data,
  cache_folder = "multiple_biomarker_groups",
  file_stem = "multiple_biomarker_groups", mcmc_pars = mcmc_pars,
  antigenic_map = multi_antigenic_map,
  possible_exposure_times = possible_exposure_times,
  data_type = c("discrete", "continuous")
)
```

```{r multiple-biomarker-output,fig.height=5,fig.width=8}
multi_fit$all_diagnostics$theta_estimates %>%
  filter(names %in% c("boost_short", "obs_sd", "min_measurement",
                      "max_measurement")) %>%
  select(names, median, lower95_CrI, upper95_CrI)

plot_estimated_antibody_model(
  multi_fit$mcmc_chains$theta_chain, settings = multi_fit$settings,
  solve_times = possible_exposure_times, by_group = TRUE
) + coord_cartesian(ylim = c(0, 10))
```

Model fits for biomarker group 1 observations.
```{r multiple-biomarker-output1}
## Biomarker group 1
plot_model_fits(
  multi_fit$mcmc_chains$theta_chain, multi_fit$mcmc_chains$inf_chain,
  known_infection_history = multi_simulated$infection_histories,
  settings = multi_fit$settings, individuals = 1:3,
  orientation = "cross-sectional", expand_to_all_times = TRUE
)[[1]]
```

Model fits for biomarker group 2 observations.
```{r multiple-biomarker-output2}
## Biomarker group 2
plot_model_fits(
  multi_fit$mcmc_chains$theta_chain, multi_fit$mcmc_chains$inf_chain,
  known_infection_history = multi_simulated$infection_histories,
  settings = multi_fit$settings, individuals = 1:3,
  orientation = "cross-sectional", expand_to_all_times = TRUE
)[[2]]
```
# Prior version 1

The infection-history prior controls how the model describes the probability of
infection across the possible exposure times and how new histories are
proposed during MCMC. In the default `prior_version = 2`, the infection-history
prior is described using the beta-distributed infection-rate parameters in
`par_tab`; the model then integrates over the corresponding time-specific
probabilities.

`prior_version = 1` instead has a separate `phi` parameter for every possible
exposure time. At a given parameter draw, `phi` is directly the probability of
infection at that time. This gives more direct control over the assumed
infection pattern: the starting value, bounds, and prior can vary by time, and
individual probabilities can be fixed when they are known or assumed.
`infection_model_prior_shape1` and `infection_model_prior_shape2` are still
required entries in the parameter table, but they do not provide the
time-specific infection probabilities in this version.

This gives two related but different ways to describe infection rates. Under
prior version 1, each `phi` is an estimated parameter representing the
probability of infection in one particular exposure period, conditional on an
individual being alive and at risk. These parameters can therefore be read as
time-specific modelled attack rates, with a posterior distribution for each
time. They are probabilities used to describe the infection-history prior;
they are not the realised fraction infected in this finite simulated sample.

The attack rates shown by `plot_attack_rates()` are calculated differently.
For each posterior draw, the function counts the inferred infection states at
each time and divides by the number of people at risk. This is the posterior
attack rate implied by the latent infection histories. It is the main attack
rate summary under prior version 2, where the infection-history prior is
controlled by beta-distributed parameters rather than a separate `phi` for
each time. The same infection-history summary can also be plotted for a
prior-version-1 fit, but it should not be confused with the posterior density
of the `phi` parameters.

Prior version 2 is the default because the per-time `phi` parameters are
strongly correlated with the latent infection states. For example, a higher
value of `phi` at a particular time can often be offset by changing which
individuals are assigned an infection at that time. Sampling both quantities
directly therefore makes the MCMC explore the posterior inefficiently and can
lead to slow mixing. Prior version 2 avoids this extra set of sampled
parameters by integrating over the time-specific infection probabilities under
their beta prior when evaluating the infection-history prior. Prior version 1
is useful when those probabilities need to be controlled separately, but is
usually more difficult to fit.

```{r prior-version-1-setup,fig.height=4,fig.width=7}
prior1_par_tab <- check_par_tab(
  base_par_tab, mcmc = TRUE, version = 1,
  possible_exposure_times = possible_exposure_times, verbose = FALSE
)

phi_rows <- which(prior1_par_tab$names == "phi")
true_phi <- simulated_data$attack_rates %>%
  filter(population_group == 1) %>%
  arrange(time) %>%
  pull(prob_infection)
stopifnot(length(true_phi) == length(phi_rows))

## Use the generating probabilities rather than the realised attack rates.
## The latter are random sample proportions from the simulated histories.
phi_start_values <- true_phi
prior1_par_tab$values[phi_rows] <- phi_start_values
prior1_par_tab$lower_start[phi_rows] <- pmax(phi_start_values - 0.05, 0.001)
prior1_par_tab$upper_start[phi_rows] <- pmin(phi_start_values + 0.05, 0.999)

## Fix two time-specific infection probabilities.
fixed_phi_times <- c(1, 25)
fixed_phi_rows <- phi_rows[match(fixed_phi_times, possible_exposure_times)]
prior1_par_tab$fixed[fixed_phi_rows] <- 1

## These are separate, time-specific priors on phi, not the prior-version-2
## infection_model_prior_shape1 and infection_model_prior_shape2 entries.
phi_prior_concentration <- rep(10, length(phi_rows))
phi_prior_concentration[13] <- 100
phi_prior_shape1 <- pmax(phi_start_values * phi_prior_concentration, 0.5)
phi_prior_shape2 <- pmax((1 - phi_start_values) * phi_prior_concentration, 0.5)

prior1_prior_func <- function(par_tab) {
  phi_rows <- which(par_tab$names == "phi")
  phi_prior_concentration <- rep(10, length(phi_rows))
  phi_prior_concentration[13] <- 100
  phi_start_values <- par_tab$values[phi_rows]
  phi_prior_shape1 <- pmax(phi_start_values * phi_prior_concentration, 0.5)
  phi_prior_shape2 <- pmax((1 - phi_start_values) * phi_prior_concentration, 0.5)
  function(pars) {
    sum(dbeta(
      pars[phi_rows], phi_prior_shape1, phi_prior_shape2, log = TRUE
    ))
  }
}

phi_table <- prior1_par_tab %>%
  filter(names == "phi") %>%
  transmute(
    time = possible_exposure_times,
    probability = values,
    fixed = fixed,
    prior_mean = phi_prior_shape1 / (phi_prior_shape1 + phi_prior_shape2),
    prior_concentration = phi_prior_concentration
  )

## Show the attack-rate distributions implied by three of the phi priors.
n_alive <- get_n_alive(antibody_data, possible_exposure_times)
prior_prediction_times <- c(5, 13, 20)
set.seed(103)
prior_attack_rate_draws <- bind_rows(lapply(prior_prediction_times, function(i) {
  phi_draws <- rbeta(10000, phi_prior_shape1[i], phi_prior_shape2[i])
  data.frame(
    time = possible_exposure_times[i],
    attack_rate = rbinom(length(phi_draws), n_alive[i], phi_draws) / n_alive[i]
  )
}))
ggplot(prior_attack_rate_draws, aes(attack_rate)) +
  geom_histogram(bins = 50, fill = "grey70", colour = "white") +
  facet_wrap(~time) +
  labs(x = "Attack rate in one exposure period", y = "Prior predictive draws") +
  theme_classic()
```

```{r prior-version-1-call, eval=FALSE}
prior1_fit <- serosolver(
  par_tab = prior1_par_tab, antibody_data = antibody_data,
  antigenic_map = antigenic_map,
  possible_exposure_times = possible_exposure_times,
  prior_version = 1, prior_func = prior1_prior_func,
  filename = "prior_version_1", data_type = "discrete",
  n_chains = 3, parallel = TRUE, mcmc_pars = mcmc_pars
)
```

```{r prior-version-1-fit, include=FALSE}
prior1_fit <- fit_or_load(
  par_tab = prior1_par_tab, antibody_data = antibody_data,
  cache_folder = "prior_version_1", file_stem = "prior_version_1",
  mcmc_pars = mcmc_pars, antigenic_map = antigenic_map,
  possible_exposure_times = possible_exposure_times,
  prior_version = 1, prior_func = prior1_prior_func, data_type = "discrete"
)
```

The posterior densities of the `phi` parameters show the estimated
time-specific infection probabilities directly. The fixed `phi` values are
shown in `phi_table` above rather than as densities, since they are point
values rather than estimated quantities.

```{r prior-version-1-phi-density}
phi_chain_names <- make.unique(as.character(prior1_par_tab$names))
phi_chain_names <- phi_chain_names[
  prior1_par_tab$names == "phi" & prior1_par_tab$fixed == 0
]

prior1_fit$mcmc_chains$theta_chain %>%
  select(samp_no, chain_no, all_of(phi_chain_names)) %>%
  pivot_longer(-c(samp_no, chain_no), names_to = "parameter") %>%
  mutate(parameter = factor(parameter, levels = phi_chain_names)) %>%
  ggplot(aes(value, fill = factor(chain_no))) +
  geom_density(alpha = 0.25) +
  facet_wrap(~parameter, scales = "free") +
  labs(x = "Per-time probability of infection", y = "Posterior density",
       fill = "Chain") +
  theme_classic()
```

```{r prior-version-1-output}
phi_estimates <- prior1_fit$mcmc_chains$theta_chain %>%
  select(all_of(phi_chain_names)) %>%
  pivot_longer(everything(), names_to = "parameter") %>%
  group_by(parameter) %>%
  summarise(
    median = median(value),
    lower = quantile(value, 0.025),
    upper = quantile(value, 0.975),
    .groups = "drop"
  ) %>%
  mutate(time = possible_exposure_times[match(parameter, phi_chain_names)])

true_attack_rates <- simulated_data$attack_rates %>%
  transmute(population_group, time, AR) %>%
  filter(is.finite(AR))
true_phi_rates <- simulated_data$attack_rates %>%
  transmute(population_group, time, AR = prob_infection)

attack_rate_plot <- plot_attack_rates(
  prior1_fit$mcmc_chains$inf_chain, settings = prior1_fit$settings,
  by_group = FALSE, plot_den = FALSE, pad_chain = TRUE,
  true_ar = NULL
) +
  labs(colour = "Estimated attack rate from inferred infection histories\n(colour indicates samples taken)") +
  guides(
    colour = guide_legend(nrow = 2),
    shape = guide_legend(nrow = 2)
  )

attack_rate_plot +
  geom_point(
    data = true_attack_rates,
    aes(
      x = time - 0.18, y = AR,
      shape = "True attack rate from simulated infection histories"
    ),
    inherit.aes = FALSE, colour = "black", size = 2.5
  ) +
  geom_pointrange(
    data = phi_estimates,
    aes(
      x = time + 0.18, y = median, ymin = lower, ymax = upper,
      shape = "Estimated phi"
    ),
    inherit.aes = FALSE, colour = "black", size = 0.4
  ) +
  geom_point(
    data = true_phi_rates,
    aes(x = time + 0.18, y = AR, shape = "True phi used in simulation"),
    inherit.aes = FALSE, colour = "black", size = 2
  ) +
  scale_shape_manual(
    name = "Reference and phi quantities",
    values = c(
      "True attack rate from simulated infection histories" = 1,
      "Estimated phi" = 16,
      "True phi used in simulation" = 2
    )
  )
```

# References and related material

The [serosolver guide](serosolver-guide.html) covers the main workflow, MCMC
file handling, and general diagnostics. The [demographic variables and
covariates vignette](demographics_covariates.html) covers fixed and
time-varying covariates and the separate variant-specific workaround. The
case-study vignettes provide longer examples using real datasets.
