---
title: "Case study 2: cross-sectional data"
subtitle: Inferring historical A/H3N2 attack rates using cross-sectional serology
csl: plos.csl
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
editor_options:
  chunk_output_type: console
bibliography: vignette_references.bib
vignette: |
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{Serosolver Case Study 2: Fluscape}
  %\VignetteEngine{knitr::rmarkdown}
---

<!-- Modified by an AI assistant on 2026-09-17 using GPT-5. Updated worked
observation-model examples to use character data_type labels. -->
<!-- Modified by an AI assistant on 2026-09-16 using GPT-5. Updated simulation-recovery fitting to use the base parameter table while preserving the existing vignette text and code comments. -->

```{r setup, include=FALSE}
## Keep the simulated data reproducible when the MCMC caches are regenerated.
set.seed(1)

knitr::opts_chunk$set(
  collapse = TRUE,
  echo = TRUE,
  comment = "#>",
  eval = TRUE,
  fig.width = 8,
  fig.height = 6,
  dpi = 300,
  out.width = "100%"
)
options(digits = 3)

## Keep case-study inputs and MCMC caches relative to this vignette.
vignette_dir <- dirname(knitr::current_input())
case_study_dir <- file.path(vignette_dir, "..", "inst", "extdata", "case_study_2")
```

# Overview
This vignette demonstrates how to use `serosolver` to infer antibody kinetics and historical attack rates from cross-sectional haemagglutination inhibition (HI) titre data measured against multiple influenza A/H3N2 strains. Unlike Case Study 1, where longitudinal samples provide information about changes in antibody levels over time, each individual here has one sampling time. The antibody measurements against multiple, antigenically related viruses provide information about the timing of previous infections.

The data are from a serosurvey in Guangzhou, China, collected in 2009 [@kucharski2015]. The analysis uses an annual infection-history resolution and an antigenic map based on Fonville et al. The model estimates long-term antibody kinetics and annual attack rates while fixing the short-term kinetics parameters, which are not well identified by these cross-sectional data alone.

The vignette covers data preparation, antigenic-map construction, model assumptions and priors, fitting and checking the model, and simulation recovery. MCMC chains are written to the Case Study 2 folder when a fit is needed; subsequent renders read those chains instead of refitting the model.

# Setup

```{r packages,warning=FALSE,message=FALSE}
## Install serosolver from GitHub when needed.
## remotes::install_github("seroanalytics/serosolver")

library(serosolver)
library(ggplot2)
library(dplyr)
library(reshape2)
library(doParallel)
library(coda)
```

# Preparing the model inputs
## Serological data
The data used in this analysis are haemagglutination inhibition (HI) titres against multiple influenza A/H3N2 strains, representing viruses which have circulated since A/H3N2 first emerged in 1968. All samples were taken in 2009, and infections are inferred at an annual resolution. The raw data are in a wide format, providing the highest two-fold dilution of serum at which haemagglutination is inhibited.

```{r data-preparation,warning=FALSE,message=FALSE}
sample_year <- 2009

## Read in the raw wide-format data.
raw_dat_path <- file.path(case_study_dir, "Fluscape_HI_data.csv")
raw_dat <- read.csv(raw_dat_path, stringsAsFactors = FALSE)

## Add an index for each individual before converting to long format.
raw_dat$individual <- seq_len(nrow(raw_dat))

## Convert the titre data to long format.
melted_dat <- reshape2::melt(
  raw_dat,
  id.vars = c("individual", "Age"),
  variable.name = "biomarker_id",
  value.name = "measurement"
)

## Extract the year represented by each virus column.
melted_dat$biomarker_id <- as.numeric(
  sub("HI.H3N2.", "", melted_dat$biomarker_id, fixed = TRUE)
)

## Remove missing measurements and convert HI titres to log2 units.
melted_dat <- melted_dat[complete.cases(melted_dat), ]
melted_dat$measurement[melted_dat$measurement == 0] <- 5
melted_dat$measurement <- log2(melted_dat$measurement / 5)

## Convert age at sampling to birth time, and record the common sample time.
antibody_data <- melted_dat %>%
  transmute(
    individual,
    birth = sample_year - Age,
    sample_time = sample_year,
    biomarker_id,
    measurement,
    repeat_number = 1L,
    biomarker_group = 1L
  )

## Validate the antibody data before fitting.
antibody_data <- check_data(antibody_data)
```

The resulting data contain one observation for each measured individual-virus combination. The measurement scale is log2 HI titre relative to a titre of 5, so the lowest recorded titre becomes zero. All samples were collected in 2009; the possible infection times run from the emergence of A/H3N2 in 1968 through the sample year.

```{r data-summary}
## Summarise the measurements and the observed virus panel.
range(antibody_data$measurement)
table(antibody_data$biomarker_id)
```

```{r data-plot}
## View the data in the format used by the model.
plot_antibody_data(
  antibody_data,
  possible_exposure_times = 1968:2009,
  n_indivs = 16,
  study_design = "cross-sectional"
) + facet_wrap(~ individual, ncol = 4)
```

Given that this analysis uses titres from multiple, antigenically related viruses, it is necessary to define an antigenic map describing the antigenic distance between all of the viruses here. We use coordinates based on the antigenic map created by Fonville et al. Generating the antigenic map involves fitting a smoothing spline through provided coordinates to give a representative virus for each time point (in this case, each year) that an individual could be infected. This also supplies antigenic coordinates for years for which no measured virus is available.

The point labels show the year of circulation for each virus, with the antigenic coordinates interpolated from the provided data.

```{r antigenic-map}
## Read the approximate antigenic coordinates.
antigenic_coords_path <- file.path(case_study_dir, "fonville_map_approx.csv")
antigenic_coords <- read.csv(antigenic_coords_path, stringsAsFactors = FALSE)

virus_key <- c(
  "HK68" = 1968, "EN72" = 1972, "VI75" = 1975, "TX77" = 1977,
  "BK79" = 1979, "SI87" = 1987, "BE89" = 1989, "BJ89" = 1989,
  "BE92" = 1992, "WU95" = 1995, "SY97" = 1997, "FU02" = 2002,
  "CA04" = 2004, "WI05" = 2005, "PE06" = 2006
)
antigenic_coords$Strain <- unname(virus_key[antigenic_coords$Strain])
antigenic_map <- generate_antigenic_map_flexible(antigenic_coords)
antigenic_map <- antigenic_map[antigenic_map$inf_times <= sample_year, ]
possible_exposure_times <- antigenic_map$inf_times

ggplot(antigenic_map) +
  geom_point(aes(x = x_coord, y = y_coord)) +
  geom_text(aes(x = x_coord, y = y_coord, label = inf_times), vjust = -0.5) +
  theme_minimal() +
  labs(title = "Antigenic map", x = "Antigenic coordinate 1", y = "Antigenic coordinate 2")
```

## Model assumptions
The parameter table controls which parameters are estimated or fixed, their starting values, and their allowable ranges. With one sample per individual, the short-term antibody response is not well separated from the timing of infection. We therefore estimate long-term boosting, cross-reactivity, antigenic seniority, and observation error, while fixing the short-term parameters at zero. This case study uses the default linear waning form; exponential waning is selected through a fixed `exponential_waning` row in `par_tab`, as described in the [advanced features vignette](ADVANCED_FEATURES_VIGNETTE_LINK).

```{r model-setup}
## The same starting parameter table is used in Case Study 1.
data(example_par_tab)
par_tab <- example_par_tab

## Estimate long-term parameters and observation error.
par_tab[par_tab$names %in% c("boost_long", "cr_long", "antigenic_seniority", "obs_sd"), "fixed"] <- 0

## Fix short-term kinetics parameters, which are not well identified here.
short_term_pars <- c("boost_short", "wane_short", "cr_short", "wane_long")
par_tab[par_tab$names %in% short_term_pars, "fixed"] <- 1
par_tab[par_tab$names %in% short_term_pars, "values"] <- 0

## Set the delay from infection to peak response to zero for the annual model.
par_tab[par_tab$names == "boost_delay", "values"] <- 0

## Match the bounded observation model to the transformed data.
par_tab[par_tab$names == "min_measurement", c("values", "lower_bound", "upper_bound")] <- 0
par_tab[par_tab$names == "max_measurement", c("values", "lower_bound", "upper_bound")] <- max(antibody_data$measurement)

## Check the parameter table before fitting.
par_tab <- check_par_tab(
  par_tab,
  mcmc = TRUE,
  possible_exposure_times = possible_exposure_times,
  version = 2
)
```

The assumed antibody kinetics can be inspected before fitting:

```{r antibody-model}
plot_antibody_model(
  pars = par_tab,
  infection_history = possible_exposure_times[c(1, 3, 13, 18, 28, 38)],
  antigenic_map = antigenic_map,
  times = possible_exposure_times
)[[2]]
```


## Priors
`serosolver` is a Bayesian model, so prior distributions must be specified for the model parameters. Prior predictive output helps assess whether these assumptions produce plausible antibody responses and infection rates before the data are used.

The prior function below follows the weak regularising priors used in Case Study 1: standard normal priors for any stratification coefficients, a log-normal prior for the antibody boost, weak Beta priors for bounded parameters, and a broad log-normal prior for observation error. The infection-history prior is set to a modal annual attack rate of around 15%, with relatively high variance.

```{r priors}
## Set the prior mode for annual attack rates.
ar_priors <- find_beta_prior_mode(0.15, 4)
par_tab[par_tab$names == "infection_model_prior_shape1", "values"] <- ar_priors$shape1
par_tab[par_tab$names == "infection_model_prior_shape2", "values"] <- ar_priors$shape2

## Define priors for the free antibody-model parameters.
prior_func <- function(par_tab) {
  par_names <- par_tab$names
  coef_pars <- which(grepl("coef", par_names))

  function(pars) {
    names(pars) <- par_names
    ## Weak prior for any stratification coefficients, as in Case Study 1.
    prior_p <- sum(dnorm(pars[coef_pars], 0, 1, log = TRUE))

    ## Weak priors for the free antibody-model parameters.
    prior_p <- prior_p + dlnorm(pars["boost_long"], log(2), 0.5, log = TRUE)
    prior_p <- prior_p + dbeta(pars["cr_long"], 1, 10, log = TRUE)
    prior_p <- prior_p + dbeta(pars["antigenic_seniority"], 1, 10, log = TRUE)
    prior_p <- prior_p + dlnorm(pars["obs_sd"], log(1), 0.5, log = TRUE)
    prior_p
  }
}
```

```{r prior-predictive, eval=TRUE, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
## Run the prior predictive model only when its cached chains are absent.
prior_dir <- file.path(case_study_dir, "prior_mcmc_chains")
prior_file <- file.path(prior_dir, "prior")
prior_n_chains <- 3
prior_mcmc_pars <- c(
  adaptive_iterations = 20000,
  iterations = 50000,
  thin = 10,
  thin_inf_hist = 100,
  save_block = 1000
)
dir.create(prior_dir, recursive = TRUE, showWarnings = FALSE)

prior_settings_file <- paste0(prior_file, "_serosolver_settings.RData")
prior_chain_files <- list.files(
  prior_dir,
  pattern = "^prior_[0-9]+_chain\\.csv$",
  full.names = TRUE
)
prior_history_files <- sub("_chain\\.csv$", "_infection_histories.csv", prior_chain_files)
prior_cache_available <- length(prior_chain_files) == prior_n_chains &&
  file.exists(prior_settings_file) &&
  all(file.exists(prior_history_files))

if (!prior_cache_available) {
  prior_res <- serosolver(
    par_tab = par_tab,
    antibody_data = antibody_data,
    possible_exposure_times = possible_exposure_times,
    filename = prior_file,
    prior_func = prior_func,
    data_type = "discrete",
    solve_likelihood = FALSE,
    n_chains = prior_n_chains,
    parallel = TRUE,
    mcmc_pars = prior_mcmc_pars,
    verbose = FALSE
  )
} else {
  load(prior_settings_file)
  prior_settings <- serosolver_settings
  prior_chains <- load_mcmc_chains(
    location = prior_dir,
    par_tab = par_tab,
    burnin = prior_mcmc_pars[["adaptive_iterations"]],
    thin = prior_mcmc_pars[["thin"]],
    convert_mcmc = FALSE,
    verbose = FALSE
  )
  prior_res <- list(
    all_diagnostics = plot_mcmc_diagnostics(
      location = prior_dir,
      par_tab = par_tab,
      burnin = prior_mcmc_pars[["adaptive_iterations"]]
    ),
    settings = prior_settings,
    plot_antibody_model = plot_estimated_antibody_model(
      prior_chains$theta_chain,
      settings = prior_settings,
      solve_times = possible_exposure_times
    ),
    mcmc_chains = prior_chains
  )
}
```

The prior predictive antibody-model plot is available from the returned object:

```{r prior-output}
prior_res$plot_antibody_model + coord_cartesian(ylim=c(0,8))
```

## Stratifying model parameters
No demographic or covariate stratification is used in this case study. The
multiple-virus measurements provide the cross-sectional information used to
infer infection timing.

# Fitting `serosolver`

The main `serosolver()` call uses the antibody data, antigenic map, parameter
table, and possible exposure times to estimate the posterior distribution of
the antibody-model parameters and infection histories. The MCMC chains are
saved to disk together with a progress text file and settings file.

```{r fit-model-call, eval=FALSE}
mcmc_pars <- c(
  adaptive_iterations = 100000,
  iterations = 200000,
  thin = 100,
  thin_inf_hist = 1000,
  save_block = 1000,
  proposal_inf_hist_time_prop = 1,
  proposal_inf_hist_indiv_prop = 1,
  proposal_inf_hist_group_swap_ratio = 0.8,
  proposal_inf_hist_group_swap_prop = 1
)

res <- serosolver(
  par_tab = par_tab,
  antibody_data = antibody_data,
  antigenic_map = antigenic_map,
  possible_exposure_times = possible_exposure_times,
  filename = mcmc_file,
  prior_func = prior_func,
  data_type = "discrete",
  n_chains = 3,
  parallel = TRUE,
  mcmc_pars = mcmc_pars,
  verbose = TRUE
)
```

```{r fit-model, eval=TRUE, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
## Fit the model once, then use the saved chains for subsequent renders.
mcmc_dir <- file.path(case_study_dir, "fluscape_mcmc_chains")
mcmc_file <- file.path(mcmc_dir, "fluscape")
mcmc_n_chains <- 3
dir.create(mcmc_dir, recursive = TRUE, showWarnings = FALSE)

mcmc_pars <- c(
  adaptive_iterations = 100000,
  iterations = 200000,
  thin = 100,
  thin_inf_hist = 1000,
  save_block = 1000,
  proposal_inf_hist_time_prop = 1,
  proposal_inf_hist_indiv_prop = 1,
  proposal_inf_hist_group_swap_ratio = 0.8,
  proposal_inf_hist_group_swap_prop = 1
)

mcmc_settings_file <- paste0(mcmc_file, "_serosolver_settings.RData")
mcmc_chain_files <- list.files(
  mcmc_dir,
  pattern = "^fluscape_[0-9]+_chain\\.csv$",
  full.names = TRUE
)
mcmc_history_files <- sub("_chain\\.csv$", "_infection_histories.csv", mcmc_chain_files)
mcmc_cache_available <- length(mcmc_chain_files) == mcmc_n_chains &&
  file.exists(mcmc_settings_file) &&
  all(file.exists(mcmc_history_files))

if (!mcmc_cache_available) {
  res <- serosolver(
    par_tab = par_tab,
    antibody_data = antibody_data,
    antigenic_map = antigenic_map,
    possible_exposure_times = possible_exposure_times,
    filename = mcmc_file,
    prior_func = prior_func,
    data_type = "discrete",
    n_chains = mcmc_n_chains,
    parallel = TRUE,
    mcmc_pars = mcmc_pars,
    verbose = FALSE
  )
} else {
  load(mcmc_settings_file)
  fit_settings <- serosolver_settings
  chains <- load_mcmc_chains(
    location = mcmc_dir,
    par_tab = par_tab,
    burnin = mcmc_pars[["adaptive_iterations"]],
    thin = mcmc_pars[["thin"]],
    convert_mcmc = FALSE,
    verbose = FALSE
  )
  max_samps <- min(mcmc_pars[["iterations"]] / mcmc_pars[["thin"]], nrow(chains$theta_chain))
  res <- list(
    all_diagnostics = plot_mcmc_diagnostics(
      location = mcmc_dir,
      par_tab = par_tab,
      burnin = mcmc_pars[["adaptive_iterations"]]
    ),
    settings = fit_settings,
    antibody_predictions = plot_antibody_predictions(
      chains$theta_chain,
      chains$inf_chain,
      settings = fit_settings,
      nsamp = max_samps
    ),
    plot_fits_cross_sectional = plot_model_fits(
      chains$theta_chain,
      chains$inf_chain,
      individuals = 1:min(5, length(unique(antibody_data$individual))),
      settings = fit_settings,
      orientation = "cross-sectional",
      expand_to_all_times = FALSE,
      expand_to_all_biomarker_ids = TRUE
    ),
    plot_attack_rates = plot_attack_rates(
      chains$inf_chain,
      settings = fit_settings,
      plot_den = FALSE
    ),
    plot_antibody_model = plot_estimated_antibody_model(
      chains$theta_chain,
      settings = fit_settings,
      solve_times = possible_exposure_times
    ),
    mcmc_chains = chains
  )
}
```

```{r load-chains, eval=TRUE, include=FALSE}
chains <- res$mcmc_chains
serosolver_settings <- res$settings
```

MCMC diagnostics assess whether the chains have converged to a common posterior
distribution and whether the effective sample size is sufficient after
accounting for autocorrelation. The criteria used here follow typical Bayesian
modelling practice, although the less efficient sampler used by `serosolver`
means that we generally use somewhat more relaxed criteria for $\hat{R}$ and
effective sample size than would be expected from a Stan model. The [Stan
reference manual](https://mc-stan.org/docs/reference-manual/analysis.html) is
a useful general resource.

```{r diagnostics}
## MCMC convergence diagnostics.
print(res$all_diagnostics$p_thetas[[1]] +
        ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, hjust = 1)))
print(res$all_diagnostics$p_thetas[[2]] +
        ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, hjust = 1)))
knitr::kable(res$all_diagnostics$theta_estimates)
knitr::kable(head(res$all_diagnostics$inf_hist_estimates$by_year))
```

The model-fit plots compare observed titres with posterior predictions for
individuals. The orange regions show posterior uncertainty about infection
times, while the titre predictions show whether the antibody model describes
the observed cross-sectional measurements.

```{r automatic-plots1}
res$antibody_predictions$p_pointrange
res$antibody_predictions$p_hist_draws
res$antibody_predictions$proportion_correct
```


```{r automatic-plots2,fig.width=7,fig.height=4}
res$plot_fits_cross_sectional[[1]] + facet_wrap(~individual,ncol=3)
```

The same individual-level plot can be recreated manually from the saved
chains. This is useful when changing the individuals or number of posterior
draws without refitting the model.

```{r manual-model-fits}
plot_model_fits(
  chain = chains$theta_chain,
  infection_histories = chains$inf_chain,
  individuals = 1:9,
  orientation = "cross-sectional",
  settings = serosolver_settings,
  expand_to_all_times = FALSE,
  expand_to_all_biomarker_ids = TRUE
)[[1]] + facet_wrap(~ individual, ncol = 3)
```

## Model estimates

The main estimates are the antibody kinetics, historical attack rates, and
posterior infection timings.

```{r model-estimates1,fig.height=4,fig.width=7}
res$plot_attack_rates
```


```{r model-estimates2}
res$plot_antibody_model
res$all_diagnostics$p_inf_hists$indiv_infections
```

```{r infection-estimates}
plot_cumulative_infection_histories(
  inf_chain = chains$inf_chain,
  indivs = 1:25,
  possible_exposure_times = possible_exposure_times
)[[2]] +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, hjust = 1))
```

The infection-history chain can also be summarised by age at the possible time
of infection. This describes inferred infection frequency across age groups.

```{r age-infection-summary}
chains$inf_chain %>%
  as.data.frame() %>%
  rename(individual = i, time = j, infection = x) %>%
  mutate(time = possible_exposure_times[time]) %>%
  left_join(antibody_data %>% select(individual, birth) %>% distinct(), by = "individual") %>%
  mutate(age = time - birth) %>%
  filter(age >= 0) %>%
  mutate(age_group = age %/% 10) %>%
  group_by(individual, age_group, samp_no, chain_no) %>%
  summarise(infection_freq = sum(infection), .groups = "drop") %>%
  group_by(individual, age_group) %>%
  summarise(mean_inf_freq = mean(infection_freq), .groups = "drop") %>%
  ggplot() +
  geom_boxplot(aes(x = age_group, y = mean_inf_freq, group = age_group)) +
  xlab("Age at possible infection") +
  ylab("Mean number of infections") +
  theme_minimal()
```

# Simulation recovery

Simulation recovery checks whether the inference machinery can recover known
parameter values. Data are simulated from the model using known values and
then refitted using the same workflow as the real data.

```{r simulation-setup}
## Use the fitted parameter estimates to define the simulated antibody model.
best_pars <- get_best_pars(chains$theta_chain)
best_pars <- best_pars[names(best_pars) != "total_infections"]
sim_par_tab <- par_tab
sim_par_tab$values <- best_pars[match(sim_par_tab$names, names(best_pars))]

## Simulate annual attack rates and a cross-sectional serosurvey.
sim_attack_rates <- simulate_attack_rates(
  infection_years = possible_exposure_times,
  mean_par = 0.15,
  sd_par = 0.25,
  n_groups = 1
)

simulated_data <- simulate_data(
  par_tab = sim_par_tab,
  n_indiv = nrow(antibody_data %>% select(individual) %>% distinct()),
  antigenic_map = antigenic_map,
  sampling_times = sample_year,
  nsamps = 1,
  attack_rates = sim_attack_rates,
  data_type = "discrete",
  measured_biomarker_ids = unique(antibody_data$biomarker_id)
)

plot_antibody_data(
  simulated_data$antibody_data,
  possible_exposure_times = possible_exposure_times,
  n_indivs = 16,
  study_design = "cross-sectional",
  infection_histories = simulated_data$infection_histories
) + facet_wrap(~ individual, ncol = 4)
```

The simulated data are fitted with shorter MCMC settings for this vignette.
The cache is used automatically after the first successful run.

```{r simulation-fit-call, eval=FALSE}
sim_res <- serosolver(
  par_tab = sim_par_tab,
  antibody_data = simulated_data$antibody_data,
  antigenic_map = antigenic_map,
  possible_exposure_times = possible_exposure_times,
  filename = sim_mcmc_file,
  prior_func = prior_func,
  data_type = "discrete",
  n_chains = 3,
  parallel = TRUE,
  mcmc_pars = sim_mcmc_pars,
  verbose = TRUE
)
```

```{r simulation-fit, eval=TRUE, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
sim_mcmc_dir <- file.path(case_study_dir, "simulation_mcmc_chains")
sim_mcmc_file <- file.path(sim_mcmc_dir, "simulated")
sim_n_chains <- 3
dir.create(sim_mcmc_dir, recursive = TRUE, showWarnings = FALSE)

sim_mcmc_pars <- c(
  adaptive_iterations = 10000,
  iterations = 20000,
  thin = 10,
  thin_inf_hist = 100,
  save_block = 1000
)

sim_fit_par_tab <- sim_par_tab
sim_settings_file <- paste0(sim_mcmc_file, "_serosolver_settings.RData")
sim_chain_files <- list.files(
  sim_mcmc_dir,
  pattern = "^simulated_[0-9]+_chain\\.csv$",
  full.names = TRUE
)
sim_history_files <- sub("_chain\\.csv$", "_infection_histories.csv", sim_chain_files)
sim_cache_available <- length(sim_chain_files) == sim_n_chains &&
  file.exists(sim_settings_file) &&
  all(file.exists(sim_history_files))

if (!sim_cache_available) {
  sim_res <- serosolver(
    par_tab = sim_fit_par_tab,
    antibody_data = simulated_data$antibody_data,
    antigenic_map = antigenic_map,
    possible_exposure_times = possible_exposure_times,
    filename = sim_mcmc_file,
    prior_func = prior_func,
    data_type = "discrete",
    n_chains = sim_n_chains,
    parallel = TRUE,
    mcmc_pars = sim_mcmc_pars,
    verbose = FALSE
  )
} else {
  load(sim_settings_file)
  sim_settings <- serosolver_settings
  sim_chains <- load_mcmc_chains(
    location = sim_mcmc_dir,
    par_tab = sim_fit_par_tab,
    burnin = sim_mcmc_pars[["adaptive_iterations"]],
    thin = sim_mcmc_pars[["thin"]],
    convert_mcmc = FALSE,
    verbose = FALSE
  )
  sim_res <- list(
    all_diagnostics = plot_mcmc_diagnostics(
      location = sim_mcmc_dir,
      par_tab = sim_fit_par_tab,
      burnin = sim_mcmc_pars[["adaptive_iterations"]]
    ),
    settings = sim_settings,
    mcmc_chains = sim_chains
  )
}
sim_chains <- sim_res$mcmc_chains
```

The parameter estimates can be compared with the values used to simulate the
data. Recovery is good when posterior medians are close to the true values and
the 95% credible intervals generally include them.

```{r simulation-parameter-recovery}
sim_res$all_diagnostics$theta_estimates %>%
  filter(names != "total_infections") %>%
  left_join(
    sim_fit_par_tab %>% select(names, true = values),
    by = "names"
  ) %>%
  select(names, true, median, lower95_CrI, upper95_CrI) %>%
  knitr::kable(digits = 3)
```

The attack-rate estimates should be close to the simulated attack rates:

```{r simulation-attack-rates}
plot_attack_rates(
  sim_chains$inf_chain,
  settings = sim_res$settings,
  true_ar = simulated_data$attack_rates,
  plot_den = FALSE
)
```

The cumulative infection-history plot gives a second check on the timing of
infections. Dashed lines show the true cumulative number of infections for
each simulated individual; these should generally lie within the orange
posterior intervals.

```{r simulation-cumulative-infections,fig.width=8,fig.height=6}
plot_cumulative_infection_histories(
  inf_chain = sim_chains$inf_chain,
  indivs = 1:9,
  real_inf_hist = simulated_data$infection_histories,
  possible_exposure_times = possible_exposure_times,
  number_col = 3,
  pad_chain = TRUE
)[[1]] +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, hjust = 1))
```

Finally, compare the fitted model with the simulated observations and true
infection histories. In this plot, dashed lines show the true infection times.
They should generally pass through dark orange regions, which indicate a high
posterior probability of infection.

```{r simulation-model-fits}
plot_model_fits(
  sim_chains$theta_chain,
  sim_chains$inf_chain,
  known_infection_history = simulated_data$infection_histories,
  individuals = 1:9,
  orientation = "cross-sectional",
  settings = sim_res$settings,
  expand_to_all_times = FALSE,
  expand_to_all_biomarker_ids = TRUE,
  p_ncol = 1
)[[1]] + facet_wrap(~ individual, ncol = 3)
```
