---
title: "Demographic variables and covariates"
subtitle: Using fixed and time-varying demographic information in `serosolver`
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
editor_options:
  chunk_output_type: console
bibliography: vignette_references.bib
vignette: |
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{Demographic variables and covariates}
  %\VignetteEngine{knitr::rmarkdown}
---

<!-- Modified by an AI assistant on 2026-09-17 using GPT-5. Updated worked
observation-model examples to use character data_type labels. -->
<!-- Modified by an AI assistant on 2026-09-16 using GPT-5. Restored the requested variant-stratification warning and removed unnecessary compact data checks. -->

```{r setup, include=FALSE}
set.seed(1)
knitr::opts_chunk$set(
  collapse = TRUE, echo = TRUE, comment = "#>", eval = TRUE,
  fig.width = 8, fig.height = 6, dpi = 300, out.width = "100%"
)
options(digits = 3)
vignette_dir <- dirname(knitr::current_input())
cache_dir <- file.path(vignette_dir, "..", "inst", "extdata", "demographics_covariates")
loader_file <- file.path(vignette_dir, "..", "inst", "extdata",
                         "vignette_chain_loader.R")
if (!file.exists(loader_file)) {
  loader_file <- system.file("extdata", "vignette_chain_loader.R",
                             package = "serosolver")
}
source(loader_file)
demographic_bundles <- load_vignette_chain_bundle(
  vignette_dir, "demographics_covariates.rds"
)
```

# Overview

Demographic and other covariate information is useful when infection rates or
antibody kinetics are expected to differ between groups. `serosolver` treats
these as separate choices. `population_group` defines groups with different
infection-history priors and is used when summarising attack rates. The
`stratification` column in `par_tab` defines demographic effects on
antibody-model parameters such as antibody boosting or waning rates.

You can mix and match different covariates to affect different parameters. For example, location could define
different infection rates by location, while vaccination status and age group could affect antibody
kinetics. `serosolver` has a flexible interface for specifying these covariates. For example, fixed covariates can be included as variables in `antibody_data` or in a
separate one-row-per-individual `demographics` table. Users can also allow covariate membership to vary over time using the
demographics table with one row for each individual and possible exposure time combination, which is a suitable approach for covariates which change with time, such as treatment group or age.

This vignette uses a synthetic longitudinal dataset with continuous antibody measurements, five samples per individual, 25 possible
exposure times, two infection-rate groups, fixed vaccination and location covariates, and a time-varying age group. Variant-specific parameters are introduced in a separate major section below; other extensions are covered in the advanced-features vignette.

When known stratification effects are needed for a simulation-recovery check,
they can be supplied to `simulate_data()` in a `coefficient_values` data
frame. Each row gives the base parameter, stratification variable, non-baseline
level, biomarker group, and coefficient value. The same base `par_tab` is then
used for fitting; `serosolver()` creates the coefficient rows from its
`stratification` entries.

## Package setup

```{r packages, message=FALSE, warning=FALSE}
## Install serosolver from GitHub when needed.
## remotes::install_github("seroanalytics/serosolver")
library(serosolver)
library(ggplot2)
library(dplyr)
library(tidyr)
library(doParallel)
library(coda)
```

# Covariate stratification

## Create and inspect the synthetic data

```{r create-data}
set.seed(123)

n_indiv <- 100
possible_exposure_times <- 1:25
sample_times <- c(5, 8, 11, 14, 17)

## Covariate factors are zero-based; population_group IDs are one-based.
## Create a version of the demographics data frame which has fixed covariate membership
fixed_demographics <- data.frame(
  individual = seq_len(n_indiv),
  birth = sample(1:5, n_indiv, replace = TRUE),
  vaccinated = rep(0:1, length.out = n_indiv),
  location = rep(0:1, each = n_indiv / 2)
) %>%
  mutate(population_group = location + 1L)

## Create a version of the demographics data frame with time-varying covariate membership
timevarying_demographics <- tidyr::expand_grid(
  fixed_demographics, time = possible_exposure_times
) %>%
  arrange(individual, time) %>%
  mutate(age_group = as.integer(time - birth >= 5))

## Use one stratified table both to simulate and to fit the data.
data(example_par_tab)
par_tab <- example_par_tab
par_tab$stratification <- NA_character_

## Simplify the kinetics model
par_tab[par_tab$names %in% c("cr_long", "cr_short", "wane_long"), "fixed"] <- 1
par_tab[par_tab$names %in% c("cr_long", "cr_short", "wane_long"), "values"] <- 0
par_tab[par_tab$names == "min_measurement", c("values", "lower_bound", "upper_bound")] <- 0
par_tab[par_tab$names == "max_measurement", c("values", "lower_bound", "upper_bound")] <- 10

## Stratify attack rates by population group
par_tab[par_tab$names %in% c(
  "infection_model_prior_shape1", "infection_model_prior_shape2"
), "stratification"] <- "population_group"

## Stratify short-term boosting magnitude by age group and vaccination status
par_tab[par_tab$names == "boost_short", "stratification"] <- "age_group, vaccinated"

## Set known, well-separated coefficient values for the simulation. The
## coefficient table uses the same parameter and biomarker-group labels as
## par_tab, but does not need the generated coefficient names.
coefficient_values <- data.frame(
  parameter = c("boost_short", "boost_short"),
  stratification = c("age_group", "vaccinated"),
  stratification_level = c(1, 1),
  biomarker_group = c(1, 1),
  value = c(1, -1)
)

sim_attack_rates <- simulate_attack_rates(
  possible_exposure_times, mean_par = c(0.05, 0.35),
  sd_par = c(0.05, 0.05), n_groups = 2
)

## Continuous observations use data_type = "continuous".
simulated_data <- simulate_data(
  par_tab = par_tab, n_indiv = n_indiv,
  possible_exposure_times = possible_exposure_times,
  measured_biomarker_ids = 25, sampling_times = sample_times,
  nsamps = length(sample_times), demographics = timevarying_demographics,
  attack_rates = sim_attack_rates, data_type = "continuous",
  coefficient_values = coefficient_values, verbose = FALSE
)
antibody_data <- check_data(simulated_data$antibody_data)
```


```{r create-data1, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
## Use the generated coefficient rows when summarising model outputs.
diagnostic_par_tab <- simulated_data$par_tab
fit_par_tab <- par_tab
```


```{r data-plot}
plot_antibody_data(
  antibody_data, possible_exposure_times = possible_exposure_times,
  n_indivs = 12, study_design = "longitudinal"
) +
  facet_wrap(~ individual, ncol = 4) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Ways to pass fixed covariates

Fixed demographic values can be included directly in `antibody_data` when
they are constant for each individual.

```{r fixed-in-antibody-data}
fixed_covariates <- fixed_demographics %>%
  select(individual, vaccinated, location, population_group)
antibody_data_with_fixed_covariates <- antibody_data %>%
  left_join(fixed_covariates, by = "individual")
head(check_data(antibody_data_with_fixed_covariates))
```

Alternatively, use a separate demographics table to pass any covariate information separately.

```{r fixed-demographics}
fixed_demographics_for_model <- fixed_demographics %>%
  select(individual, birth, vaccinated, location, population_group)
head(check_demographics(fixed_demographics_for_model))
```

## Time-varying demographic variables

A time-varying table has `individual`, `birth`, and `time` columns at a minimum,
together with the covariates that change over time. Here, `age_group` is
calculated from birth and model time, as individuals change age groups as time progresses. In this version, every individual must have a row for
every possible exposure time.

```{r time-varying-demographics}
head(check_demographics(timevarying_demographics))
```

## Set up `par_tab`

`par_tab` is the model control table. In this example, the infection-history prior
parameters are stratified by the `population_group` ID, while
`boost_short` has additive antibody-kinetics effects for `age_group`
and `vaccinated`.

```{r model-setup}
check_par_tab(
  par_tab, mcmc = TRUE,
  possible_exposure_times = possible_exposure_times, version = 2
)
par_tab %>%
  filter(!is.na(stratification)) %>%
  select(names, stratification)
simulated_data$par_tab %>%
  filter(par_type == 4) %>%
  select(names, values, par_type, biomarker_group)
fixed_demographics %>%
  select(location, population_group) %>%
  distinct() %>%
  arrange(location)
```

Covariate levels used for parameter stratification are zero-based, with level
0 as the baseline. `population_group` is different: it is one-based. Here,
location 0 maps to population group 1 and location 1 maps to population group 2.

## How covariate coefficients are used

When a parameter is stratified, the value in its original `par_tab` row is the
value for the level-0 baseline group. `serosolver` adds a `par_type = 4`
coefficient row for each additional level. For example, a coefficient for
`boost_short` and `vaccinated = 1` describes the change from the
`vaccinated = 0` value; it is not a second independent value for
`boost_short`.

The coefficient is combined with the baseline value on a scale that keeps the
parameter within its allowed range. If the parameter is unbounded, the value
for group `g` is

\[
x_g = x_0 + \beta_g,
\]

where `x_0` is the baseline value and `beta_g` is the coefficient. For a
positive parameter such as `boost_short`, the coefficient is added on the log
scale:

\[
x_g = \exp\{\log(x_0) + \beta_g\}
    = x_0\exp(\beta_g).
\]

For a parameter bounded between 0 and 1, it is added on the logit scale:

\[
x_g = \operatorname{logit}^{-1}
       \{\operatorname{logit}(x_0) + \beta_g\}.
\]

When more than one covariate is listed in `stratification`, the corresponding
coefficients are added together on the relevant transformed scale. Every
coefficient row added in this way must therefore be included in the prior
function used by the Bayesian model.

For example, with age group `a` and vaccination status `v`, the combined
effect can be written as

\[
h(x_{a,v}) = h(x_0) + \beta^{\mathrm{age}}_a
                    + \beta^{\mathrm{vaccinated}}_v,
\]

where `h` is the identity, log, or logit transformation described above.
The natural-scale value is obtained by applying the inverse transformation
after the coefficients have been added.

Baseline antibody parameters describe the group coded 0 for each stratifying
covariate. A coefficient describes the change relative to that baseline.
`population_group` selects a group-specific infection-history prior; it is not
an antibody-kinetics coefficient. Sparse groups can produce weakly identified
coefficients, so the prior may contribute substantial uncertainty to the
implied group differences.

## Priors for stratification coefficients

Every coefficient row generated from `par_tab$stratification` is an estimated
parameter and needs a prior. The following function identifies the added
coefficient rows by their names and applies a weak standard normal prior to
them. It is passed to both the prior-predictive and fitted `serosolver()` calls
below.

```{r prior-function}
prior_func <- function(par_tab) {
  par_names <- par_tab$names
  coef_pars <- which(grepl("coef", par_names))
  function(pars) {
    names(pars) <- par_names
    ## Place normal prior on the stratification coefficients
    prior_p <- sum(dnorm(pars[coef_pars], 0, 1, log = TRUE))
    prior_p <- prior_p + dlnorm(pars["boost_long"], log(2), 0.5, log = TRUE)
    prior_p <- prior_p + dlnorm(pars["boost_short"], log(2), 0.5, log = TRUE)
    prior_p <- prior_p + dbeta(pars["wane_short"], 4, 8, log = TRUE)
    prior_p <- prior_p + dlnorm(pars["obs_sd"], log(1), 0.5, log = TRUE)
    prior_p
  }
}

simulated_data$par_tab %>%
  filter(par_type == 4) %>%
  select(names, values)
```

A standard normal prior on a coefficient for a positive parameter such as
`boost_short` corresponds to a multiplicative effect on the natural scale.

```{r coefficient-prior,fig.height=4,fig.width=6}
coefficient_prior <- tibble(
  coefficient = rnorm(10000), multiplier = exp(coefficient)
)
ggplot(coefficient_prior, aes(multiplier)) +
  geom_histogram(bins = 60, fill = "grey70", colour = "white") +
  labs(x = "Multiplicative effect relative to baseline", y = "Prior draws") +
  theme_classic()
```

## Combining covariates

Several covariates can be listed in one `stratification` entry, separated by a
comma. For example, `age_group, vaccinated` tells `serosolver` to estimate one
set of age-group coefficients and one set of vaccination coefficients for the
parameter. These effects are added together on the appropriate transformed
scale, so this does not automatically create a separate parameter for every
age and vaccination combination. For example, the effect for an older,
vaccinated individual is the baseline value plus the age effect plus the
vaccination effect.

The infection-history priors are grouped using `population_group`, which is a
single group identifier for each individual. If infection rates should differ
according to a combination of fixed covariates, create one `population_group`
column representing those combined groups before fitting. The mapping between
the original covariates and the resulting group IDs should then be checked
before the model is run.

```{r combined-covariates}
fixed_demographics %>%
  count(population_group, location, vaccinated) %>%
  arrange(population_group, location, vaccinated)
par_tab %>%
  filter(names == "boost_short" | grepl("infection_model_prior", names)) %>%
  select(names, stratification)
```

# Fit and inspect the covariate-stratified model

The complete call below uses separate time-varying demographics, continuous
observations, and the explicit `prior_func`. It is shown without evaluation
so the full workflow is visible.

```{r fit-model-call, eval=FALSE}
mcmc_pars <- c(
  adaptive_iterations = 5000, iterations = 10000, thin = 20,
  thin_inf_hist = 100, save_block = 1000
)
res <- serosolver(
  par_tab = fit_par_tab, antibody_data = antibody_data,
  demographics = timevarying_demographics,
  possible_exposure_times = possible_exposure_times,
  filename = "demographic_fit",
  prior_func = prior_func, data_type = "continuous", n_chains = 3,
  parallel = TRUE, mcmc_pars = mcmc_pars, verbose = TRUE
)
```

```{r prior-predictive, include=FALSE}
prior_dir <- file.path(cache_dir, "prior_mcmc_chains")
prior_file <- file.path(prior_dir, "prior_covariates")
if (!is.null(demographic_bundles[["prior"]])) {
  prior_dir <- materialize_vignette_chain_fit(
    demographic_bundles[["prior"]], tempfile("serosolver_demographic_prior_"),
    "prior_covariates"
  )
}
prior_mcmc_pars <- c(
  adaptive_iterations = 5000, iterations = 10000, thin = 20,
  thin_inf_hist = 100, save_block = 1000
)
dir.create(prior_dir, recursive = TRUE, showWarnings = FALSE)
prior_settings_file <- paste0(prior_file, "_serosolver_settings.RData")
prior_chain_files <- list.files(
  prior_dir, pattern = "^prior_covariates_[0-9]+_chain\\.csv$", full.names = TRUE
)
prior_history_files <- sub(
  "_chain\\.csv$", "_infection_histories.csv", prior_chain_files
)
prior_cache_available <- length(prior_chain_files) == 3L &&
  file.exists(prior_settings_file) && all(file.exists(prior_history_files))

if (!prior_cache_available) {
  prior_res <- serosolver(
    par_tab = fit_par_tab, antibody_data = antibody_data,
    demographics = timevarying_demographics,
    possible_exposure_times = possible_exposure_times,
    filename = prior_file, prior_func = prior_func, data_type = "continuous",
    solve_likelihood = FALSE, n_chains = 3, parallel = TRUE,
    mcmc_pars = prior_mcmc_pars, verbose = FALSE
  )
} else {
  load(prior_settings_file)
  prior_settings <- serosolver_settings
  prior_chains <- load_mcmc_chains(
    location = prior_dir, par_tab = fit_par_tab,
    burnin = prior_mcmc_pars[["adaptive_iterations"]],
    thin = 1, convert_mcmc = FALSE,
    verbose = FALSE
  )
  prior_res <- list(
    all_diagnostics = plot_mcmc_diagnostics(
      location = prior_dir, par_tab = diagnostic_par_tab,
      burnin = prior_mcmc_pars[["adaptive_iterations"]]
    ),
    settings = prior_settings,
    mcmc_chains = prior_chains
  )
}
```

```{r fit-model, include=FALSE}
mcmc_dir <- file.path(cache_dir, "mcmc_chains")
mcmc_file <- file.path(mcmc_dir, "demographic_fit")
if (!is.null(demographic_bundles[["main"]])) {
  mcmc_dir <- materialize_vignette_chain_fit(
    demographic_bundles[["main"]], tempfile("serosolver_demographic_main_"),
    "demographic_fit"
  )
  mcmc_file <- file.path(mcmc_dir, "demographic_fit")
}
mcmc_pars <- c(
  adaptive_iterations = 5000, iterations = 10000, thin = 20,
  thin_inf_hist = 100, save_block = 1000
)
dir.create(mcmc_dir, recursive = TRUE, showWarnings = FALSE)
mcmc_settings_file <- paste0(mcmc_file, "_serosolver_settings.RData")
mcmc_chain_files <- list.files(
  mcmc_dir, pattern = "^demographic_fit_[0-9]+_chain\\.csv$", full.names = TRUE
)
mcmc_history_files <- sub(
  "_chain\\.csv$", "_infection_histories.csv", mcmc_chain_files
)
mcmc_cache_available <- length(mcmc_chain_files) == 3L &&
  file.exists(mcmc_settings_file) && all(file.exists(mcmc_history_files))

if (!mcmc_cache_available) {
  res <- serosolver(
    par_tab = fit_par_tab, antibody_data = antibody_data,
    demographics = timevarying_demographics,
    possible_exposure_times = possible_exposure_times,
    filename = mcmc_file, prior_func = prior_func, data_type = "continuous",
    n_chains = 3, parallel = TRUE, mcmc_pars = mcmc_pars, verbose = FALSE
  )
} else {
  load(mcmc_settings_file)
  fit_settings <- serosolver_settings
  chains <- load_mcmc_chains(
    location = mcmc_dir, par_tab = fit_par_tab,
    burnin = mcmc_pars[["adaptive_iterations"]],
    thin = 1, convert_mcmc = FALSE, verbose = FALSE
  )
  max_samps <- nrow(chains$theta_chain)
  res <- list(
    all_diagnostics = plot_mcmc_diagnostics(
      location = mcmc_dir, par_tab = diagnostic_par_tab,
      burnin = mcmc_pars[["adaptive_iterations"]]
    ),
    settings = fit_settings,
    antibody_predictions = plot_antibody_predictions(
      chains$theta_chain, chains$inf_chain, settings = fit_settings,
      nsamp = max_samps
    ),
    plot_attack_rates = plot_attack_rates(
      chains$inf_chain, settings = fit_settings,
      by_group = TRUE, plot_den = FALSE
    ),
    plot_antibody_model = plot_estimated_antibody_model(
      chains$theta_chain, settings = fit_settings,
      solve_times = possible_exposure_times, data_type = "continuous"
    ),
    mcmc_chains = chains
  )
}
chains <- res$mcmc_chains
serosolver_settings <- res$settings
```

The prior-predictive results below show the range of values implied by the
chosen priors for the main antibody parameters.

```{r prior-output}
prior_res$all_diagnostics$theta_estimates %>%
  filter(names %in% c("boost_long", "boost_short", "wane_short")) %>%
  select(names, median, lower95_CrI, upper95_CrI)
```

The fitted attack rates are compared with the known simulated values. The
points and intervals are inferred from the antibody data; the simulated
values provide a recovery check rather than additional fitted data.

```{r attack-rate-output}
plot_attack_rates(
  chains$inf_chain,
  true_ar = simulated_data$attack_rates %>% filter(is.finite(AR)),
  settings = serosolver_settings, by_group = TRUE, plot_den = FALSE
) + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r antibody-and-coefficients}
res$plot_antibody_model + coord_cartesian(ylim = c(0, 10))

coefficient_estimates <- res$all_diagnostics$theta_estimates %>%
  filter(grepl("coef", names)) %>%
  select(names, median, lower95_CrI, upper95_CrI, ess)
coefficient_truth <- simulated_data$par_tab %>%
  filter(par_type == 4) %>% select(names, true_value = values)
left_join(coefficient_estimates, coefficient_truth, by = "names")
```

The fitted kinetics can also be shown with the demographic groups overlaid in
each panel. Set `by_group = FALSE` to
colour the group-specific trajectories in the same panel, or `by_group = TRUE` to plot them as separate panels.

```{r antibody-model-overlaid}
plot_estimated_antibody_model(
  chains$theta_chain, settings = serosolver_settings,
  solve_times = possible_exposure_times, data_type = "continuous", by_group = FALSE
) + coord_cartesian(ylim = c(0, 10))
```

# Variant-stratified parameters

> **Warning:** Variant stratification is not a generally supported package
> workflow. This implementation is a project-specific workaround used solely
> for the analysis in [McCormack *et al.* [@mccormack2026]](https://www.medrxiv.org/content/10.64898/2026.04.16.26351029v1).
> Hence, the example code below is fairly customised and complicated. You have been warned!

The demographic examples above allow antibody parameters to differ according
to who was infected, or according to a person's demographic characteristics at
the time of infection. A related but different question is whether the
antibody response depends on which variant or strain caused the infection.
This is useful when the same biomarker can be boosted by infections with
several variants, and the model should estimate different kinetics for those
variants.

Variant stratification is not a demographic grouping of individuals. Instead,
it attaches a group to each possible infection time in the `antigenic_map`.
The `exposure_group` column identifies the variant associated with that time,
so the model can use the appropriate variant-specific parameters when an
infection is assigned to that time. This example leaves demographic effects
out of the variant fit so that the variant-specific parameterisation is shown
on its own.

This section uses the same number of synthetic individuals and sample times as
the first analysis, but does not use demographic or covariate stratification.
It creates two antibody-parameter blocks with different short-term boosting
values, and fits the simulated data to test whether those differences can be
recovered.

The implementation of variant stratification currently reuses the
`biomarker_group` structure that normally represents different observation
types. This is a crude workaround developed for a specific project. It means
that variant stratification cannot currently be combined with genuinely
multiple observed biomarker groups in the same model. The code below therefore
temporarily creates two groups for `simulate_data()`, then retains only group 1
as the observed data used for fitting. The second group is not a second
biomarker measurement in this analysis.

There are three different group labels in this workflow. `biomarker_group` is
normally used to identify separate observed data types, such as antibody and
avidity measurements. In this variant-specific workaround it is instead
reused in `par_tab` and the antigenic map to label separate
antibody-parameter blocks. `exposure_group` is attached to the possible
infection times in `antigenic_map` and tells the model which parameter block to
use for an infection at that time. For example, if times 1--10 have
`exposure_group = 1` and times 11--25 have `exposure_group = 2`, an infection
assigned to time 15 uses the parameters for exposure group 2. The duplicated
map rows provide a corresponding map block for each parameter block; they do
not represent separate observed biomarker measurements in the fitted model.

`population_group` has a different role. It is attached to people (or to
people at particular times when demographics vary over time) and selects the
infection-history prior used for that person. It is used for group-specific
infection histories and attack-rate summaries; it does not identify the
variant that caused an infection or select the variant-specific antibody
parameters.
In this example everyone is assigned to population group 1, so there is no
demographic stratification in the variant fit.

## Set up the variant map and parameter blocks

```{r variant-simulation, results='hide'}
variant_map_one_group <- data.frame(
  x_coord = if_else(
    possible_exposure_times <= 12,
    possible_exposure_times / 12,
    2 + possible_exposure_times / 12
  ),
  y_coord = if_else(possible_exposure_times <= 12, 0, 2),
  inf_times = possible_exposure_times,
  exposure_group = if_else(possible_exposure_times <= 12, 1L, 2L),
  biomarker_group = 1L
)

## The antigenic map has one copy for each antibody-parameter block.
variant_antigenic_map <- bind_rows(
  variant_map_one_group,
  variant_map_one_group %>% mutate(biomarker_group = 2L)
)

## Use an unstratified table for the variant-specific example.
variant_base_par_tab <- par_tab
variant_base_par_tab$stratification <- NA_character_

## Duplicate the base antibody-parameter rows for the second variant block.
variant_par_rows <- variant_base_par_tab %>%
  filter(!names %in% c(
    "infection_model_prior_shape1", "infection_model_prior_shape2",
    "exponential_waning"
  )) %>%
  mutate(biomarker_group = 2L)
variant_par_tab <- bind_rows(variant_base_par_tab, variant_par_rows)
variant_par_tab[
  variant_par_tab$biomarker_group == 2 & variant_par_tab$names == "boost_short",
  "values"
] <- 3.5

variant_fit_par_tab <- variant_par_tab
variant_demographics <- timevarying_demographics %>%
  select(individual, birth, time) %>%
  mutate(population_group = 1L)
variant_sim_attack_rates <- simulate_attack_rates(
  possible_exposure_times, mean_par = 0.2, sd_par = 0.05, n_groups = 1
)

variant_simulated_data <- simulate_data(
  par_tab = variant_par_tab, n_indiv = n_indiv,
  antigenic_map = variant_antigenic_map,
  possible_exposure_times = possible_exposure_times,
  measured_biomarker_ids = possible_exposure_times,
  sampling_times = sample_times, nsamps = length(sample_times),
  demographics = variant_demographics,
  attack_rates = variant_sim_attack_rates,
  data_type = c("continuous", "continuous"),
  verbose = FALSE
)
## simulate_data creates both groups because variant_par_tab contains both
## parameter blocks. Retain one observed group for the fitted model.
variant_antibody_data <- variant_simulated_data$antibody_data %>%
  filter(biomarker_group == 1) %>%
  check_data()
```

The map assigns exposure groups to infection times rather than to people.
In this workaround, `biomarker_group` labels the variant-specific parameter
blocks rather than two distinct observed antibody measurements. Because
`simulate_data()` uses the same labels when constructing simulated data, it
initially returns rows for both labels. Only group 1 is retained as the single
observed biomarker group for fitting; group 2 is the duplicated simulation
block needed to implement the variant-specific parameterisation.

```{r variant-map,fig.height=4,fig.width=6}
ggplot(variant_map_one_group, aes(x_coord, y_coord, colour = factor(exposure_group))) +
  geom_point(size = 2) +
  labs(colour = "Exposure group", x = "Antigenic x coordinate", y = "Antigenic y coordinate") +
  theme_classic()
```

```{r variant-parameter-blocks}
variant_par_tab %>%
  filter(names == "boost_short") %>%
  select(names, biomarker_group, values)
variant_antibody_data %>%
  count(biomarker_group, sample_time) %>%
  summarise(n_groups = n_distinct(biomarker_group),
            n_samples = n_distinct(sample_time), .groups = "drop")
```

The parameter values used to generate the two blocks are deliberately
different. This makes the simulation-recovery comparison informative rather
than testing two nearly identical models.

```{r variant-prior-function, include=FALSE}
variant_prior_func <- function(par_tab) {
  par_names <- par_tab$names
  boost_long_pars <- which(par_names == "boost_long")
  boost_short_pars <- which(par_names == "boost_short")
  wane_short_pars <- which(par_names == "wane_short")
  obs_sd_pars <- which(par_names == "obs_sd")
  function(pars) {
    names(pars) <- par_names
    prior_p <- sum(dlnorm(pars[boost_long_pars], log(2), 0.5, log = TRUE))
    prior_p <- prior_p + sum(dlnorm(pars[boost_short_pars], log(2), 0.5, log = TRUE))
    prior_p <- prior_p + sum(dbeta(pars[wane_short_pars], 4, 8, log = TRUE))
    prior_p <- prior_p + sum(dlnorm(pars[obs_sd_pars], log(1), 0.5, log = TRUE))
    prior_p
  }
}
```

## Fit the variant-stratified model

The complete variant-stratified fit is:

```{r variant-fit-call, eval=FALSE}
variant_mcmc_pars <- c(
  adaptive_iterations = 5000, iterations = 10000, thin = 20,
  thin_inf_hist = 100, save_block = 1000
)
variant_res <- serosolver(
  par_tab = variant_fit_par_tab,
  antibody_data = variant_antibody_data,
  demographics = variant_demographics,
  antigenic_map = variant_antigenic_map,
  possible_exposure_times = possible_exposure_times,
  filename = "variant_fit",
  prior_func = variant_prior_func,
  data_type = "continuous", n_chains = 3, parallel = TRUE,
  mcmc_pars = variant_mcmc_pars, verbose = TRUE
)
```

```{r variant-fit, include=FALSE, message=FALSE, warning=FALSE}
variant_mcmc_dir <- file.path(cache_dir, "variant_mcmc_chains")
variant_mcmc_file <- file.path(variant_mcmc_dir, "variant_fit")
if (!is.null(demographic_bundles[["variant"]])) {
  variant_mcmc_dir <- materialize_vignette_chain_fit(
    demographic_bundles[["variant"]], tempfile("serosolver_demographic_variant_"),
    "variant_fit"
  )
  variant_mcmc_file <- file.path(variant_mcmc_dir, "variant_fit")
}
variant_mcmc_pars <- c(
  adaptive_iterations = 5000, iterations = 10000, thin = 20,
  thin_inf_hist = 100, save_block = 1000
)
dir.create(variant_mcmc_dir, recursive = TRUE, showWarnings = FALSE)
variant_settings_file <- paste0(variant_mcmc_file, "_serosolver_settings.RData")
variant_chain_files <- list.files(
  variant_mcmc_dir, pattern = "^variant_fit_[0-9]+_chain\\.csv$",
  full.names = TRUE
)
variant_history_files <- sub(
  "_chain\\.csv$", "_infection_histories.csv", variant_chain_files
)
variant_fit_available <- length(variant_chain_files) == 3L &&
  file.exists(variant_settings_file) && all(file.exists(variant_history_files))

if (variant_fit_available) {
  load(variant_settings_file)
  variant_fit_available <- length(unique(
    serosolver_settings$antibody_data$biomarker_group
  )) == 1L
}

if (!variant_fit_available) {
  variant_res <- serosolver(
    par_tab = variant_fit_par_tab,
    antibody_data = variant_antibody_data,
    demographics = variant_demographics,
    antigenic_map = variant_antigenic_map,
    possible_exposure_times = possible_exposure_times,
    filename = variant_mcmc_file,
    prior_func = variant_prior_func,
    data_type = "continuous", n_chains = 3, parallel = TRUE,
    mcmc_pars = variant_mcmc_pars, verbose = FALSE
  )
} else {
  variant_settings <- serosolver_settings
  variant_chains <- load_mcmc_chains(
    location = variant_mcmc_dir, par_tab = variant_fit_par_tab,
    burnin = variant_mcmc_pars[["adaptive_iterations"]], thin = 1,
    convert_mcmc = FALSE, verbose = FALSE
  )
  variant_res <- list(
    all_diagnostics = plot_mcmc_diagnostics(
      location = variant_mcmc_dir, par_tab = variant_par_tab,
      burnin = variant_mcmc_pars[["adaptive_iterations"]]
    ),
    settings = variant_settings,
    mcmc_chains = variant_chains
  )
}
variant_chains <- variant_res$mcmc_chains
variant_settings <- variant_res$settings
```

## Simulation recovery

```{r variant-attack-rates}
plot_attack_rates(
  variant_chains$inf_chain,
  true_ar = variant_simulated_data$attack_rates %>% filter(is.finite(AR)),
  settings = variant_settings, by_group = TRUE, plot_den = FALSE
) + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

The parameter table comparison focuses on the variant-specific boosting values.
The duplicated names are made unique in the same order used in the MCMC output.

```{r variant-parameter-recovery}
variant_truth <- variant_simulated_data$par_tab %>%
  filter(fixed == 0) %>%
  mutate(chain_name = make.unique(names)) %>%
  select(chain_name, names, biomarker_group, true = values)
variant_res$all_diagnostics$theta_estimates %>%
  rename(chain_name = names) %>%
  left_join(variant_truth, by = "chain_name") %>%
  filter(names == "boost_short") %>%
  select(names, biomarker_group, true, median, lower95_CrI, upper95_CrI) %>%
  knitr::kable(digits = 3)
```

The fitted chain contains two sets of antibody-kinetics parameters: one for
each variant-specific block. Because this workaround stores the second block
using the `biomarker_group` structure, its parameter names have a `.1` suffix
in the chain. The fitted antibody data still contain only one observed
biomarker group, so the code below does not add a second set of observations
or fit another model. Instead, it uses `plot_estimated_antibody_model()` to
make two model-implied trajectories. It chooses one possible infection time
from each `exposure_group`; the model then uses the corresponding fitted
parameter block. The trajectories are combined in one plot, using one
representative biomarker ID to keep the comparison readable.

```{r variant-kinetics-comparison,fig.height=4,fig.width=6}
variant_infection_times <- c(
  "Exposure group 1" = 6,
  "Exposure group 2" = 18
)
representative_biomarker_id <- max(
  as.numeric(as.character(variant_antibody_data$biomarker_id))
)

variant_kinetics <- lapply(seq_along(variant_infection_times), function(i) {
  infection_time <- variant_infection_times[i]
  plot_estimated_antibody_model(
    variant_chains$theta_chain,
    antibody_data = variant_antibody_data,
    demographics = variant_demographics,
    antigenic_map = variant_antigenic_map,
    possible_exposure_times = possible_exposure_times,
    par_tab = variant_fit_par_tab,
    solve_times = seq(infection_time, infection_time + 10, by = 0.1),
    set_infections = infection_time,
    data_type = "continuous", settings = variant_settings,
    by_group = TRUE, nsamp = 1000
  )$data %>%
    filter(as.numeric(as.character(biomarker_id)) == representative_biomarker_id) %>%
    mutate(
      time_since_infection = sample_time - infection_time,
      exposure_group = names(variant_infection_times)[i]
    )
}) %>%
  bind_rows()

ggplot(variant_kinetics,
       aes(time_since_infection, median,
           colour = exposure_group, fill = exposure_group)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2, colour = NA) +
  geom_line() +
  coord_cartesian(ylim = c(0, 10)) +
  labs(
    x = "Time since infection",
    y = "Antibody level",
    colour = "Variant exposure group",
    fill = "Variant exposure group"
  ) +
  theme_classic()
```

# References

This vignette focuses on demographic and covariate information, followed by the
project-specific variant-stratification workaround. Related material is
available in the following vignettes:

- the [serosolver guide](SEROSOLVER_GUIDE_LINK) for the main workflow and
  general model options;
- the [Hong Kong case study](CASE_STUDY_1_VIGNETTE_LINK) for a longitudinal
  example;
- the [Fluscape case study](CASE_STUDY_2_VIGNETTE_LINK) for a cross-sectional
  example; and
- the [advanced-features vignette](ADVANCED_FEATURES_VIGNETTE_LINK) for
  starting antibody levels, measurement offsets, exponential waning, and
  multiple observed biomarker groups.

<!-- TODO: replace the placeholder vignette links above with the final links. -->
